---
title: "NFL Pay-by-Play"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(ggplot2)

library(tm)
library(wordcloud)
library(stringr) 
```

```{r}
# Colors
# Black #000000
# Red #aa0200
# Gold #b3995d
# White #ffffff
```

```{r}
years <- c("2009","2010","2011","2012","2013","2014","2015","2016","2017", "2018")
wins <- c(8,6,13,11,12,8,5,2,6,4)
loss <- c(8,10,3,4,4,8,11,14,10,12)
tie <- c(0,0,0,1,0,0,0,0,0,0)

win_loss = data.frame(Year = years, Win = wins, Loss = loss, Tie = tie)

```


# Introduction

Strategy in American football is more important than perhaps the casual viewer may realize. Football is sometimes referred to as a "game of inches", which was taken from the quote “Football is a game of inches and inches make the champion”, credited to Hall of Fame coach Vince Lombardi. 

# About the Data

The dataset contains 449,371 observations across 255 variables. Each variable represents information about a given play during an NFL football game. Each record represents a specific play during a given game. The data contain information on plays from the 2009 - 2018 season. 

# Data Exploration

```{r}
# Read in the data
NFL <- read.csv("/Users/timhulak/Desktop/Syracuse/IST-719\ Information\ Visualization/data/NFL_Play_By_Play_2009-2018.csv",header = T, stringsAsFactors = F)
head(NFL)
```

```{r}
tail(NFL)
```

```{r}
# See the dimensions of the data
dim(NFL)
```


There are many statistical/probability columns in the dataset. Columns such as "total_home_comp_air_epa" (the total home team completion through the air expected points over average) do not lend themselves well to this analysis. This is because it adds too much granularity to the data and it is not clear how these figures were calculated. In addition, columns pertaining to ID numbers (such as player ID numbers) were dropped because the dataset also includes the names associated with the ID numbers. Therefore, the ID numbers are not useful and may prove redundant if they are, in fact, usable at all. This left the shape of the data as 449,371 observations of 167 varibales. 

```{r}
# make a vector of columns to be dropped
col_del <- c("no_score_prob","opp_fg_prob","opp_safety_prob","opp_td_prob","fg_prob","safety_prob","td_prob","extra_point_prob","two_point_conversion_prob","ep","epa","total_home_epa","total_away_epa","total_home_rush_epa","total_away_rush_epa","total_home_pass_epa","total_home_pass_epa","air_epa","yac_epa","comp_air_epa","comp_yac_epa","total_home_comp_air_epa","total_away_comp_air_epa","total_home_comp_yac_epa","total_away_comp_yac_epa","total_home_raw_air_epa", "total_away_raw_air_epa","total_home_raw_yac_epa","total_away_raw_yac_epa", "wp", "def_wp","home_wp","away_wp","wpa","home_wp_post","away_wp_post","total_home_rush_wpa","total_away_rush_wpa","total_home_pass_wpa","total_away_pass_wpa","air_wpa","yac_wpa","comp_air_wpa","comp_yac_wpa","total_home_comp_air_wpa","total_away_comp_air_wpa","total_home_comp_yac_wpa","total_home_raw_air_wpa","total_away_raw_air_wpa","total_home_raw_yac_wpa","total_away_raw_yac_wpa","passer_player_id","receiver_player_id","rusher_player_id","lateral_receiver_player_id","lateral_rusher_player_id","lateral_sack_player_id","interception_player_id","lateral_interception_player_id","punt_returner_player_id","lateral_punt_returner_player_id","kickoff_returner_player_id","lateral_kickoff_returner_player_id","punter_player_id","kicker_player_id","own_kickoff_recovery_player_id","blocked_player_id","tackle_for_loss_1_player_id","tackle_for_loss_2_player_id","qb_hit_1_player_id","qb_hit_2_player_id","forced_fumble_player_1_player_id","forced_fumble_player_2_player_id","solo_tackle_1_player_id","solo_tackle_2_player_id","assist_tackle_1_player_id","assist_tackle_2_player_id","assist_tackle_3_player_id","assist_tackle_4_player_id","pass_defense_1_player_id","pass_defense_2_player_id","fumbled_1_player_id","fumbled_2_player_id","fumble_recovery_1_team","fumble_recovery_1_player_id","fumble_recovery_2_player_id","penalty_player_id","total_away_pass_epa","total_away_comp_yac_wpa")
# Drop the columns
NFL <- NFL[ , !(names(NFL) %in% col_del)]
# Check the new dimensions
dim(NFL)
```




```{r}
# Sort the unique team abbreviations and count how many there are 
sort(unique(NFL$away_team))
length(unique(NFL$away_team))
```

# Initial Data Cleaning

There are 32 NFL teams. The data spans from 2009 - 2018 and shows 35 unique teams. During that time frame, 3 NFL teams changed their team abbreviation. The Chargers moved from San Diego, CA to Los Angeles, CA in 2017, therefore their team abbreviation changed from SD to LAC. In 2013, a social media campaign and fan petition saw the Jaguars change their team abbreviation from JAC to JAX. Finally, in 2016 the Rams moved from St. Louis, MO to Los Angeles, CA and their team abbreviation changed from STL to LA. The first step in cleaning the data is to convert the old abbreviations to the new abbreviations in order to analyze team performance as a whole. 


```{r}
# Combine JAC & JAX, SD & LAC, and STL to LA. 
NFL[NFL == "JAC"] <- "JAX"
NFL[NFL == "SD"] <- "LAC"
NFL[NFL == "STL"] <- "LA"
```

```{r}
# Confirm change
sort(unique(NFL$away_team))
length(unique(NFL$away_team))
```

In order to get a sense of each team's participation, the columns for game_id, home_team, and away_team were combined to create a "participation" column. Next, a column called "opponents" was created to help talley the tems that have faced one another.  

```{r}
NFL['participation'] <- paste(NFL$game_id, NFL$home_team,"Vs.", NFL$away_team)
NFL['opponents']<- paste(NFL$home_team,"Vs.", NFL$away_team)
```

Next, the game_seconds_remaining column was used to create a game_minutes_remaining column in order to better gauge the time remaining in a game. This was also done for the quarter_seconds_remaining and half_seconds_remaining.

```{r}
NFL['game_minutes_remaining'] <- round(NFL$game_seconds_remaining / 60,2)
NFL['quarter_minutes_remaining'] <- round(NFL$quarter_seconds_remaining / 60,2)
NFL['half_minutes_remaining'] <- round(NFL$half_seconds_remaining / 60,2)
```
Next, the year and month were extracted from the game_id column. 

```{r}
NFL$game_id <- as.character(NFL$game_id)
```

```{r}
NFL['year'] <- substr(NFL$game_id, start = 1, stop = 4)
NFL['month'] <- substr(NFL$game_id, start = 5, stop = 6)

```

```{r}
library(dplyr)

NFL <- NFL %>% mutate(month_abb = case_when(
  month == "01" ~ "Jan"
  , month == "02" ~ "Feb"
  , month == "03" ~ "Mar"
  , month == "04" ~ "Apr"
  , month == "05" ~ "May"
  , month == "06" ~ "Jun"
  , month == "07" ~ "Jul"
  , month == "08" ~ "Aug"
  , month == "09" ~ "Sep"
  , month == "10" ~ "Oct"
  , month == "11" ~ "Nov"
  , month == "12" ~ "Dec"
  ))

NFL <- NFL %>% mutate(home_div = case_when(
  home_team == "SEA" ~ "NFC West"
  , home_team == "SF" ~ "NFC West"
  , home_team == "LA" ~ "NFC West"
  , home_team == "ARI" ~ "NFC West"
  , home_team == "NO" ~ "NFC South"
  , home_team == "TB" ~ "NFC South"
  , home_team == "CAR" ~ "NFC South"
  , home_team == "ATL" ~ "NFC South"
  , home_team == "WAS" ~ "NFC East"
  , home_team == "NYG" ~ "NFC East"
  , home_team == "DAL" ~ "NFC East"
  , home_team == "PHI" ~ "NFC East"
  , home_team == "GB" ~ "NFC North"
  , home_team == "CHI" ~ "NFC North"
  , home_team == "MIN" ~ "NFC North"
  , home_team == "DET" ~ "NFC North"
  , home_team == "KC" ~ "AFC West"
  , home_team == "OAK" ~ "AFC West"
  , home_team == "LAC" ~ "AFC West"
  , home_team == "DEN" ~ "AFC West"
  , home_team == "TEN" ~ "AFC South"
  , home_team == "IND" ~ "AFC South"
  , home_team == "HOU" ~ "AFC South"
  , home_team == "JAX" ~ "AFC South"
  , home_team == "BUF" ~ "AFC East"
  , home_team == "MIA" ~ "AFC East"
  , home_team == "NE" ~ "AFC East"
  , home_team == "NYJ" ~ "AFC East"
  , home_team == "PIT" ~ "AFC North"
  , home_team == "BAL" ~ "AFC North"
  , home_team == "CLE" ~ "AFC North"
  , home_team == "CIN" ~ "AFC North"
  ))

NFL <- NFL %>% mutate(away_div = case_when(
  away_team == "SEA" ~ "NFC West"
  , away_team == "SF" ~ "NFC West"
  , away_team == "LA" ~ "NFC West"
  , away_team == "ARI" ~ "NFC West"
  , away_team == "NO" ~ "NFC South"
  , away_team == "TB" ~ "NFC South"
  , away_team == "CAR" ~ "NFC South"
  , away_team == "ATL" ~ "NFC South"
  , away_team == "WAS" ~ "NFC East"
  , away_team == "NYG" ~ "NFC East"
  , away_team == "DAL" ~ "NFC East"
  , away_team == "PHI" ~ "NFC East"
  , away_team == "GB" ~ "NFC North"
  , away_team == "CHI" ~ "NFC North"
  , away_team == "MIN" ~ "NFC North"
  , away_team == "DET" ~ "NFC North"
  , away_team == "KC" ~ "AFC West"
  , away_team == "OAK" ~ "AFC West"
  , away_team == "LAC" ~ "AFC West"
  , away_team == "DEN" ~ "AFC West"
  , away_team == "TEN" ~ "AFC South"
  , away_team == "IND" ~ "AFC South"
  , away_team == "HOU" ~ "AFC South"
  , away_team == "JAX" ~ "AFC South"
  , away_team == "BUF" ~ "AFC East"
  , away_team == "MIA" ~ "AFC East"
  , away_team == "NE" ~ "AFC East"
  , away_team == "NYJ" ~ "AFC East"
  , away_team == "PIT" ~ "AFC North"
  , away_team == "BAL" ~ "AFC North"
  , away_team == "CLE" ~ "AFC North"
  , away_team == "CIN" ~ "AFC North"
  ))

NFL <- NFL %>% mutate(pos_team_div = case_when(
  posteam == "SEA" ~ "NFC West"
  , posteam == "SF" ~ "NFC West"
  , posteam == "LA" ~ "NFC West"
  , posteam == "ARI" ~ "NFC West"
  , posteam == "NO" ~ "NFC South"
  , posteam == "TB" ~ "NFC South"
  , posteam == "CAR" ~ "NFC South"
  , posteam == "ATL" ~ "NFC South"
  , posteam == "WAS" ~ "NFC East"
  , posteam == "NYG" ~ "NFC East"
  , posteam == "DAL" ~ "NFC East"
  , posteam == "PHI" ~ "NFC East"
  , posteam == "GB" ~ "NFC North"
  , posteam == "CHI" ~ "NFC North"
  , posteam == "MIN" ~ "NFC North"
  , posteam == "DET" ~ "NFC North"
  , posteam == "KC" ~ "AFC West"
  , posteam == "OAK" ~ "AFC West"
  , posteam == "LAC" ~ "AFC West"
  , posteam == "DEN" ~ "AFC West"
  , posteam == "TEN" ~ "AFC South"
  , posteam == "IND" ~ "AFC South"
  , posteam == "HOU" ~ "AFC South"
  , posteam == "JAX" ~ "AFC South"
  , posteam == "BUF" ~ "AFC East"
  , posteam == "MIA" ~ "AFC East"
  , posteam == "NE" ~ "AFC East"
  , posteam == "NYJ" ~ "AFC East"
  , posteam == "PIT" ~ "AFC North"
  , posteam == "BAL" ~ "AFC North"
  , posteam == "CLE" ~ "AFC North"
  , posteam == "CIN" ~ "AFC North"
  ))

NFL <- NFL %>% mutate(def_team_div = case_when(
  defteam == "SEA" ~ "NFC West"
  , defteam == "SF" ~ "NFC West"
  , defteam == "LA" ~ "NFC West"
  , defteam == "ARI" ~ "NFC West"
  , defteam == "NO" ~ "NFC South"
  , defteam == "TB" ~ "NFC South"
  , defteam == "CAR" ~ "NFC South"
  , defteam == "ATL" ~ "NFC South"
  , defteam == "WAS" ~ "NFC East"
  , defteam == "NYG" ~ "NFC East"
  , defteam == "DAL" ~ "NFC East"
  , defteam == "PHI" ~ "NFC East"
  , defteam == "GB" ~ "NFC North"
  , defteam == "CHI" ~ "NFC North"
  , defteam == "MIN" ~ "NFC North"
  , defteam == "DET" ~ "NFC North"
  , defteam == "KC" ~ "AFC West"
  , defteam == "OAK" ~ "AFC West"
  , defteam == "LAC" ~ "AFC West"
  , defteam == "DEN" ~ "AFC West"
  , defteam == "TEN" ~ "AFC South"
  , defteam == "IND" ~ "AFC South"
  , defteam == "HOU" ~ "AFC South"
  , defteam == "JAX" ~ "AFC South"
  , defteam == "BUF" ~ "AFC East"
  , defteam == "MIA" ~ "AFC East"
  , defteam == "NE" ~ "AFC East"
  , defteam == "NYJ" ~ "AFC East"
  , defteam == "PIT" ~ "AFC North"
  , defteam == "BAL" ~ "AFC North"
  , defteam == "CLE" ~ "AFC North"
  , defteam == "CIN" ~ "AFC North"
  ))

NFL$home_confrence <- substr(NFL$home_div, start = 1, stop = 3)
NFL$away_confrence <- substr(NFL$away_div, start = 1, stop = 3)
NFL$pos_team_confrence <- substr(NFL$pos_team_div, start = 1, stop = 3)
NFL$def_team_confrence <- substr(NFL$def_team_div, start = 1, stop = 3)
```


```{r}
head(NFL)
```


```{r}
SF_Offense <- NFL[NFL$posteam == "SF",]
SF_Offense$BallSide <- "Offense"
SF_Defense <- NFL[NFL$defteam == "SF",]
SF_Defense$BallSide <- "Defence"
```

```{r}
head(SF_Offense)
```


```{r}
head(SF_Defense)
```

```{r}
library(reshape)
WL <- melt(win_loss, idvar="Year")
colnames(WL) <- c("Year","WinLoss","Freq")
```


Stacked Bar
```{r}
ggplot(WL, aes(fill=factor(WinLoss), y=Freq, x=Year)) + 
    geom_bar(position="stack", stat="identity", color = "white") +
    scale_fill_manual(values=c("#aa0200","#b3995d","gray")) + 
  theme(
  # get rid of panel grids
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change plot and panel background
  plot.background=element_rect(fill = "black"),
  panel.background = element_rect(fill = 'black'),
  # Change legend 
  legend.background = element_rect(fill = "black", color = NA),
  legend.key = element_rect(color = "white", fill = "black"),
  legend.title = element_text(color = "white"),
  legend.text = element_text(color = "white") 
  ) + ggtitle("Wins and Losses from 2009 - 2018") +
   theme(plot.title = element_text(color = "white", face = "bold.italic", size = 16)) + 
  labs(x = "Year", y = "Games") + theme(axis.title = element_text(face = "bold.italic", color = "white")) + theme(axis.text.x = element_text(face = "bold.italic", color = "white", size = 12)) + theme(axis.text.y = element_text(face = "bold.italic", color = "white", size = 12)) 

```

```{r}
head(SF_Offense)
```


Line Chart for Run vs Pass
```{r}
SF_Off_Filtered <- SF_Offense[c("year","play_type")]
SF_Off_Filtered <- na.omit(SF_Off_Filtered)
SF_Off_Filtered <- SF_Off_Filtered[(SF_Off_Filtered$play_type == "run") | (SF_Off_Filtered$play_type == "pass"), ]

x <- table(SF_Off_Filtered)
ldf <- as.data.frame(x)
head(ldf)
```




```{r}
pdf("/Users/timhulak/Desktop/Syracuse/IST-719\ Information\ Visualization/PosterFinal/line.pdf", width = 8, height = 3 )
ggplot(ldf,aes(y = Freq, x=year, group = play_type)) +
  geom_line(aes(color = play_type), size = 1.3) + 
  geom_point(aes(color = play_type),shape= 5, size = 3) + 
  scale_color_manual(values=c("#ffffff", "#b3995d")) +
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "black"),
    panel.background = element_rect(fill = 'black'),
    # Change legend 
    legend.background = element_rect(fill = "black", color = NA),
    legend.key = element_rect(color = "white", fill = "black"),
    legend.title = element_text(color = "white"),
    legend.text = element_text(color = "white") 
    ) + ggtitle("Run Play and Pass Plays from 2009 - 2018") +
     theme(plot.title = element_text(color = "white", face = "bold.italic", size = 16)) + 
    labs(x = "Year", y = "Count") + theme(axis.title = element_text(face = "bold.italic", color = "white")) + theme(axis.text.x = element_text(face = "bold.italic", color = "white", size = 12)) + theme(axis.text.y = element_text(face = "bold.italic", color = "white", size = 12)
           ) +
  ylim(0,700)
   
dev.off()    
    

```

# Penalties


```{r}
library(wordcloud2)

offensive_penalties <- SF_Offense[c("penalty_team","penalty_yards","penalty_type")]
offensive_penalties <- offensive_penalties[offensive_penalties$penalty_team == "SF",]
offensive_penalties <- na.omit(offensive_penalties)
offensive_penalties <- offensive_penalties[c("penalty_yards","penalty_type")]
offensive_penalties <- as.data.frame(apply(offensive_penalties,2,function(x)gsub('\\s+', '',x)))

x <- table(offensive_penalties$penalty_type)
opdf <- as.data.frame(x)


wordcloud2(data=opdf, color="#b3995d", backgroundColor = NA, size = .5)
 
```



```{r}
FS <- offensive_penalties[offensive_penalties$penalty_type == "FalseStart",]
length(FS$penalty_yards)
sum(as.numeric(FS$penalty_yards))


OH <- offensive_penalties[offensive_penalties$penalty_type == "OffensiveHolding",]
length(OH$penalty_yards)
sum(as.numeric(OH$penalty_yards))
mean(as.numeric(OH$penalty_yards))

sum(as.numeric(FS$penalty_yards)) + sum(as.numeric(FS$penalty_yards))
```


```{r}
defensive_penalties <- SF_Defense[c("penalty_team","penalty_yards","penalty_type")]
defensive_penalties <- defensive_penalties[defensive_penalties$penalty_team == "SF",]
defensive_penalties <- na.omit(defensive_penalties)
defensive_penalties <- defensive_penalties[c("penalty_yards","penalty_type")]
defensive_penalties <- as.data.frame(apply(defensive_penalties,2,function(x)gsub('\\s+', '',x)))


x <- table(defensive_penalties$penalty_type)
dpdf <- as.data.frame(x)

wordcloud2(data=dpdf , backgroundColor = NA, color="#b3995d", size = .3)

```

```{r}
DO <- defensive_penalties[defensive_penalties$penalty_type == "DefensiveOffside",]
length(DO$penalty_yards)
sum(as.numeric(DO$penalty_yards))


DPI <- defensive_penalties[defensive_penalties$penalty_type == "DefensivePassInterference",]
length(DPI$penalty_yards)
sum(as.numeric(DPI$penalty_yards))
mean(as.numeric(DPI$penalty_yards))

UR <- defensive_penalties[defensive_penalties$penalty_type == "UnnecessaryRoughness",]
length(UR$penalty_yards)
sum(as.numeric(UR$penalty_yards))
mean(as.numeric(UR$penalty_yards))

DH <- defensive_penalties[defensive_penalties$penalty_type == "DefensiveHolding",]
length(DH$penalty_yards)
sum(as.numeric(DH$penalty_yards))
mean(as.numeric(DH$penalty_yards))

sum(as.numeric(DO$penalty_yards)) + sum(as.numeric(DPI$penalty_yards)) + sum(as.numeric(UR$penalty_yards)) + sum(as.numeric(DH$penalty_yards))
```

```{r}
sum(as.numeric(offensive_penalties$penalty_yards)) + sum(as.numeric(defensive_penalties$penalty_yards)) 
```

Side of the field



```{r}
sideOfField = SF_Offense[c("year","pass_location","pass_length","run_gap")]
passlocation = na.omit(sideOfField[c("year","pass_location")])
passlength = na.omit(sideOfField[c("year","pass_length")])
rungap = na.omit(sideOfField[c("year","run_gap")])

PLoc <- as.data.frame(table(passlocation))
PLen <- as.data.frame(table(passlength))
Rgap <- as.data.frame(table(rungap))

PLoc2011 <- PLoc[PLoc$year == 2011,]
PLen2011 <- PLen[PLen$year == 2011,]
Rgap2011 <- Rgap[Rgap$year == 2011,]

PLoc2016 <- PLoc[PLoc$year == 2016,]
PLen2016 <- PLen[PLen$year == 2016,]
Rgap2016 <- Rgap[Rgap$year == 2016,]

```


# Colors
# Black #000000
# Red #aa0200
# Gold #b3995d
# White #ffffff

Pass Location 2016
```{r}
df <-  PLoc2016

# Compute percentages
df$fraction <- df$Freq / sum(df$Freq)

# Compute the cumulative percentages (top of each rectangle)
df$ymax <- cumsum(df$fraction)

# Compute the bottom of each rectangle
df$ymin <- c(0, head(df$ymax, n=-1))

# Compute label position
df$labelPosition <- (df$ymax + df$ymin) / 2

# Compute a good label
df$label <- paste0(df$pass_location,"\n", (round(df$Freq/sum(df$Freq),4))*100,"%")

# Make the plot
ggplot(df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=pass_location)) +
  scale_fill_manual(values=c("#ffffff","#b3995d","#000000")) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), color = "#aa0200", size=5) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") + 
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "gray"),
    panel.background = element_rect(fill = 'gray')) 

```
Pass Location 2011
```{r}
df <-  PLoc2011

# Compute percentages
df$fraction <- df$Freq / sum(df$Freq)

# Compute the cumulative percentages (top of each rectangle)
df$ymax <- cumsum(df$fraction)

# Compute the bottom of each rectangle
df$ymin <- c(0, head(df$ymax, n=-1))

# Compute label position
df$labelPosition <- (df$ymax + df$ymin) / 2

# Compute a good label
df$label <- paste0(df$pass_location,"\n", (round(df$Freq/sum(df$Freq),4))*100,"%")

# Make the plot
ggplot(df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=pass_location)) +
  scale_fill_manual(values=c("#ffffff","#b3995d","#000000")) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), color = "#aa0200", size=5) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") + 
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "gray"),
    panel.background = element_rect(fill = 'gray')) 

```


Pass Length 2016
```{r}
df <-  PLen2016

# Compute percentages
df$fraction <- df$Freq / sum(df$Freq)

# Compute the cumulative percentages (top of each rectangle)
df$ymax <- cumsum(df$fraction)

# Compute the bottom of each rectangle
df$ymin <- c(0, head(df$ymax, n=-1))

# Compute label position
df$labelPosition <- (df$ymax + df$ymin) / 2

# Compute a good label
df$label <- paste0(df$pass_length,"\n", (round(df$Freq/sum(df$Freq),4))*100,"%")

# Make the plot
ggplot(df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=pass_length)) +
  scale_fill_manual(values=c("#ffffff","#b3995d","#000000")) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), color = "#aa0200", size=5) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") + 
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "gray"),
    panel.background = element_rect(fill = 'gray'))
```


Pass Length 2011
```{r}
df <-  PLen2011

# Compute percentages
df$fraction <- df$Freq / sum(df$Freq)

# Compute the cumulative percentages (top of each rectangle)
df$ymax <- cumsum(df$fraction)

# Compute the bottom of each rectangle
df$ymin <- c(0, head(df$ymax, n=-1))

# Compute label position
df$labelPosition <- (df$ymax + df$ymin) / 2

# Compute a good label
df$label <- paste0(df$pass_length,"\n", (round(df$Freq/sum(df$Freq),4))*100,"%")

# Make the plot
ggplot(df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=pass_length)) +
  scale_fill_manual(values=c("#ffffff","#b3995d","#000000")) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), color = "#aa0200", size=5) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") + 
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "gray"),
    panel.background = element_rect(fill = 'gray')) 
```

Run Gap 2016
```{r}
df <-  Rgap2016

# Compute percentages
df$fraction <- df$Freq / sum(df$Freq)

# Compute the cumulative percentages (top of each rectangle)
df$ymax <- cumsum(df$fraction)

# Compute the bottom of each rectangle
df$ymin <- c(0, head(df$ymax, n=-1))

# Compute label position
df$labelPosition <- (df$ymax + df$ymin) / 2

# Compute a good label
df$label <- paste0(df$run_gap,"\n", (round(df$Freq/sum(df$Freq),4))*100,"%")

# Make the plot
ggplot(df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=run_gap)) +
  scale_fill_manual(values=c("#ffffff","#b3995d","#000000")) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), color = "#aa0200", size=5) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") + 
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "gray"),
    panel.background = element_rect(fill = 'gray')) 

```

Run Gap 2011
```{r}
df <-  Rgap2011

# Compute percentages
df$fraction <- df$Freq / sum(df$Freq)

# Compute the cumulative percentages (top of each rectangle)
df$ymax <- cumsum(df$fraction)

# Compute the bottom of each rectangle
df$ymin <- c(0, head(df$ymax, n=-1))

# Compute label position
df$labelPosition <- (df$ymax + df$ymin) / 2

# Compute a good label
df$label <- paste0(df$run_gap,"\n", (round(df$Freq/sum(df$Freq),4))*100,"%")

# Make the plot
ggplot(df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=run_gap)) +
  scale_fill_manual(values=c("#ffffff","#b3995d","#000000")) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), color = "#aa0200", size=5) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none") + 
  theme( 
    # get rid of panel grids
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change plot and panel background
    plot.background=element_rect(fill = "gray"),
    panel.background = element_rect(fill = 'gray')) 

```
