---
title: "IST 687 Final Project"
authors:
- Timothy Hulak
- Chanel Murray
- Lenerson Pyrrhus
- Mark Diaz
date: "`r Sys.time()`"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
# Fantastic Fourrrrrrr Bank:

## Abstract:
Fantastic Fourrrrrrr Bank management has tasked our data science team to figure out a way to determine credit card approvals, lines of credit, and APR for our credit card applicants. Further, they have tasked us with analyzing the initial applicant data as well as the application approval/denial results to ensure we are using the most efficient and equitable means to determine who is approved/denied for our credit card. 



## The Data :

* [Link to the dataset link](https://www.kaggle.com/rikdifos/credit-card-approval-prediction)
* There are 438,557 observations (rows/applications)
* There are 17 variables (columns/fields):
  + "ID" - Application ID number 
  + "CODE_GENDER" - Applicant gender (M/F)
  + "FLAG_OWN_CAR" - If the applicant owns a car (Y/N)
  + "FLAG_OWN_REALTY" - If the applicant owns realty (Y/N)
  + "CNT_CHILDREN" - How many children the applicant has
  + "AMT_INCOME_TOTAL" - Applicant's Annual income
  + "NAME_INCOME_TYPE" - Type of employment the applicant has (Student, Working, Pensioner, etc.)
  + "NAME_EDUCATION_TYPE" - Level of education
  + "NAME_FAMILY_STATUS" - Marital status (Single, Married, Widowed, etc.)
  + "NAME_HOUSING_TYPE" - Type of housing the applicant lives in (House, Apartment, With Parents, etc.)
  + "Age" - Age of the applicant
  + "DAYS_EMPLOYED" - How long the applicant has been employed 
  + "FLAG_MOBIL" - If the applicant owns a mobile phone
  + "FLAG_WORK_PHONE" - If the applicant owns a work/company phone
  + "FLAG_EMAIL" - If the applicant owns a work/company phone
  + "OCCUPATION_TYPE" - If the applicant owns an email address
  + "CNT_FAM_MEMBERS" - How many individuals live in the household
  


```{r}
# Imports
library("writexl")
library("ggplot2")
library("treemap")
library("readxl")
library("dplyr")
library("wordcloud2")
library('sqldf')
options(scipen = 999)
```

## Read in the data from our applications and preview the data
```{r}
# ----------------  Read in the data  -----------------
cc_app_data <- read.csv('application_record.csv')

head(cc_app_data)

summary(cc_app_data)
```

## Explore the Data
* Our key variables for considering credit approval, line of credit, and APR are the following:
  + Annual Income
  + Owning a car
  + Owning Realty
  + Number of Children
  + Housing type
  
* There are `r length(cc_app_data$ID)` applicants in our dataset. 
 
## Mesures of Central Tendancy and Variability

5 Number summaries of numeric variables

* Children:
  + Mean: `r mean(cc_app_data$CNT_CHILDREN)`
  + Median: `r median(cc_app_data$CNT_CHILDREN)`
  + Standard Deviation: `r sd(cc_app_data$CNT_CHILDREN)`
  + Minimum: `r min(cc_app_data$CNT_CHILDREN)`
  + Maximum: `r max(cc_app_data$CNT_CHILDREN)`
* Income:
  + Mean: `r mean(cc_app_data$AMT_INCOME_TOTAL)`
  + Median: `r median(cc_app_data$AMT_INCOME_TOTAL)`
  + Standard Deviation: `r sd(cc_app_data$AMT_INCOME_TOTAL)`
  + Minimum: `r min(cc_app_data$AMT_INCOME_TOTAL)`
  + Maximum: `r max(cc_app_data$AMT_INCOME_TOTAL)`
* Age:
  + Mean: `r mean(cc_app_data$Age)`
  + Median: `r median(cc_app_data$Age)`
  + Standard Deviation: `r sd(cc_app_data$Age)`
  + Minimum: `r min(cc_app_data$Age)`
  + Maximum:  `r max(cc_app_data$Age)`

# Data Visualizations
## Applicant Gender
```{r}

# ------------ Gender (Pie Chart) ------------ 
male <- length(which(cc_app_data$CODE_GENDER == "M"))
female <- length(which(cc_app_data$CODE_GENDER == "F"))

m_f_df <- data.frame(c("M","F"),c(male,female))
colnames(m_f_df) <- c("Sex","Count")

# Male/Female Pie chart
ggplot(m_f_df, aes(x="", y=Count, fill=Sex)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + ggtitle("Applicant Gender Ratio")

male_pct <- male/(male+female)
female_pct <- female/(male+female)
```

`r round(female_pct*100,2)`% of our applicants are female whereas `r round(male_pct*100,2)`% are male


```{r}
# Get list of unique values from the column 
housing_types <- unique(cc_app_data$NAME_HOUSING_TYPE)

# Create empty vector to hold counts of each value
housing_type_counts <- c()

# Append the counts of each unique value to the housing_type_counts vector
for (i in housing_types){
  type_count <- length(which(cc_app_data$NAME_HOUSING_TYPE == i))
  housing_type_counts <- c(housing_type_counts,type_count)
}


# ------------ Housing Type (Tree Map)  ------------ 
housing_type_data <- data.frame(housing_types,housing_type_counts)

# Housing Type Tree map
treemap(housing_type_data,
            index="housing_types",
            vSize="housing_type_counts",
            type="index",
            title = "Housing Types"
            )


```
The vast majority of our applicants (89.8%) appear to be living in a house or apartment. A distant second appears to be living with their parents (4.3%) and the remainder of housing types make up the remaining 18.5% of our data.

```{r}
cloud_count <- count(cc_app_data, OCCUPATION_TYPE)

wordcloud2(data=cloud_count, size=1.6)
```
In the word cloud above, we see that Laborers, Sales Staff, Managers, and Core Staff make up the majority of our applicant's Occupation Type. 

```{r}
# Age box plot
ggplot(cc_app_data, aes(x=Age)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + ggtitle("Applicant Age Distribution")

```
The median age of our applicants is `r median(cc_app_data$Age)` years old. 25% of our applicants are between `r min(cc_app_data$Age)` and 34 years old. 50% of our applicants are between `r min(cc_app_data$Age)` and `r median(cc_app_data$Age)` years old. 75% of our applicants are between `r min(cc_app_data$Age)` and 53 years old. There appears to be no outliers in our data, with respect to age.s

```{r}
# Own Realty Column Chart
own_realty <- length(which(cc_app_data$FLAG_OWN_REALTY == "Y"))
do_not_own_realty <- length(which(cc_app_data$FLAG_OWN_REALTY == "N"))

ggplot(cc_app_data, aes(x=as.factor(FLAG_OWN_REALTY), fill=as.factor(FLAG_OWN_REALTY) )) + 
  geom_bar( ) +
  scale_fill_hue(c = 40) +
  theme(legend.position="none") +
  ylab("Count") + 
  xlab("Own Realty") +
  ggtitle("Own Realty")
```
More than twice as many applicants own realty than those who do not; `r round((own_realty/length(cc_app_data$FLAG_OWN_REALTY))*100,2)`% own realty whereas `r round((do_not_own_realty/length(cc_app_data$FLAG_OWN_REALTY))*100,2)`% do not own realty.

## Define logic for approval and line of credit

* In order to get a CC from us, applicant needs 100%. 
  + AMT_INCOME_TOTAL (50%)
  + NAME_HOUSING_TYPE (10%)
  + FLAG_OWN_CAR (5%)
  + FLAG_OWN_REALTY (25%)
  + CNT_CHILDREN (10%)
  + 5% APR, +0.5% for every 1% under 100% (ex 99% = 5.5%)
  + 80% or above = approval, different APR (no collateral)
  + 50% - 79% 

* Line of Credit (20% of Income) seperated into Income Tiers = 
  + T1 = 0.01 
  + T2 = 0.05
  + T3 = 0.10
  + T4 = 0.15

* Approx 30% to taxes/etc, 50% to live on, 20% to our card

```{r}

# ----------------  Approval and Line of Credit Logic  -----------------

approval_check <- function(application){
  # Declare Base score of 0
  base_score <- 0
  
  # Declare Base APR
  base_APR <- 5
  additional_APR <- 0.5
  
  # Establish Tier of Income
  T1 <- 250000
  T2 <- 500000
  T3 <- 750000
  T4 <- 1000000
  
  # Establish line of credit rates 
  T1_rate <-  0.01 
  T2_rate <-  0.05
  T3_rate <-  0.10
  T4_rate <-  0.15
  
  # Pull out variables that are to be considered
  applicant_ID <- application[1]
  applicant_income <- as.numeric(application[6])
  applicant_house_type <- application[10]
  applicant_own_car <- application[3]
  applicant_own_realty <- application[4]
  applicant_count_of_children <- as.numeric(application[5])
  
  # Determine Applicant Tier and Points for Tier
  applicant_tier <- "No Tier"
  base_income <- 20000
  
  if(applicant_income <= T1 & applicant_income >= base_income){
    applicant_tier <- "T1"
    base_score <- base_score + 12.5
    
  } else if (applicant_income > T1 & applicant_income <= T2){
    applicant_tier <- "T2"
    base_score <- base_score + (12.5*2)
    
  } else if (applicant_income > T2 & applicant_income <= T3){
    applicant_tier <- "T3"
    base_score <- base_score + (12.5*3)
    
  } else if (applicant_income > T3 & applicant_income <= T4){
    applicant_tier <- "T4"
    base_score <- base_score + (12.5*4)
    
  } else {
    applicant_tier <- "Insuffienct Income"
    base_score <- 0
  }
  
  # Determine House Type and Own/Rent Score
  if (applicant_house_type == "House / apartment" & applicant_own_realty == "Y"){
    base_score <- base_score + 25 + 10
  } else if (applicant_house_type == "House / apartment" & applicant_own_realty == "N") {
    base_score <- base_score + 12.5 + 7
  } else if (applicant_house_type == "With parents" & applicant_own_realty == "Y") {
    base_score <- base_score + 25 + 7
  } else if (applicant_house_type == "With parents" & applicant_own_realty == "N") {
    base_score <- base_score + 12.5 + 5
  } else if (applicant_house_type == "Municipal apartment" & applicant_own_realty == "Y") {
    base_score <- base_score + 25 + 4
  } else if (applicant_house_type == "Municipal apartment" & applicant_own_realty == "N") {
    base_score <- base_score + 12.5 + 4
  } else if (applicant_house_type == "Rented apartment" & applicant_own_realty == "Y") {
    base_score <- base_score + 25 + 10
  } else if (applicant_house_type == "Rented apartment" & applicant_own_realty == "N") {
    base_score <- base_score + 12.5 + 6
  } else if (applicant_house_type == "Office apartment" & applicant_own_realty == "Y") {
    base_score <- base_score + 25 + 10
  } else if (applicant_house_type == "Office apartment" & applicant_own_realty == "N") {
    base_score <- base_score + 12.5 + 5
  } else if (applicant_house_type == "Co-op apartment" & applicant_own_realty == "Y") {
    base_score <- base_score + 25 + 10
  } else if (applicant_house_type == "Co-op apartment" & applicant_own_realty == "N") {
    base_score <- base_score + 20 + 6
  } else {
    base_score <- base_score + 0
  }
  
  # Determine car ownership score
  if (applicant_own_car == "Y"){
    base_score <- base_score + 5
  } else if (applicant_own_car == "N"){
    base_score <- base_score + 2.5
  } else {
    base_score <- base_score + 0
  }
  
  # Determine score for number of children
  if (applicant_count_of_children == 0){
    base_score <- base_score + 10
  } else if (applicant_count_of_children == 1){
    base_score <- base_score + 9
  } else if (applicant_count_of_children == 2){
    base_score <- base_score + 8
  } else if (applicant_count_of_children == 3){
    base_score <- base_score + 7
  } else if (applicant_count_of_children >= 4 & applicant_count_of_children <= 6){
    base_score <- base_score + 5
  } else if (applicant_count_of_children >= 7 & applicant_count_of_children <= 10){
    base_score <- base_score + 3
  } else if (applicant_count_of_children < 10){
    base_score <- base_score + 1
  }
  
  # Determine if applicant is approved 
  approval <- NA
  
  if(base_score >= 50){
    approval <- "Approved"
  } else {
    approval <- "Denied"
  }
  

  # Establish Line of Credit Amount
  line_of_credit <- 0
  
  # Determine Line of Credit
  if(applicant_tier == "T1" & approval == "Approved"){
    line_of_credit <- applicant_income*T1_rate
  } else if (applicant_tier == "T2" & approval == "Approved") {
    line_of_credit <- applicant_income*T2_rate
  } else if (applicant_tier == "T3" & approval == "Approved") {
    line_of_credit <- applicant_income*T3_rate
  } else if (applicant_tier == "T4" & approval == "Approved") {
    line_of_credit <- applicant_income*T4_rate
  } else {
    line_of_credit <- 0
  }
  
  # Set APR to 0
  applicant_APR <- 0
  
  # Determine APR
  if(base_score == 100 & approval == "Approved"){
    applicant_APR <- base_APR
  } else if (base_score < 100 & approval == "Approved"){
    APR_multiplier <- 100 - base_score
    applicant_APR <- base_APR + (APR_multiplier * additional_APR)
  }
  
  approval_vector <- NULL
  
  if(approval == "Approved"){
    approval_vector <- c(approval,line_of_credit,applicant_APR)
  } else {
    approval_vector <- c(approval,NA,NA)
  }
  
  return(approval_vector)
  

}



```

## Process the Applications 
```{r}

# # ----------------  Application Processing  -----------------
# 
# # Create Empty Vectors to hold data for new columns
# approval_col <- c()
# credit_col <- c()
# apr_col <- c()
# 
# 
# # Loop through rows and determine approval, credit line, and APR
# for(i in 1:nrow(cc_app_data)){
#   application_data <- as.vector(as.character(cc_app_data[i,]))
#   result <- approval_check(application_data)
#   approval_col <- c(approval_col,result[1])
#   credit_col <- c(credit_col,result[2])
#   apr_col <- c(apr_col,result[3])
#   
# }
# 
# # Save data to new DF
# application_results <- cc_app_data
# 
# # Add new columns with results
# application_results$Approval_Status <- approval_col
# application_results$Line_of_Credit <- as.numeric(credit_col)
# application_results$APR <- as.numeric(apr_col)/100
# 
# # Preview data
# head(application_results)
# 
# 
# write_xlsx(application_results, "application_results.xlsx")
```

## Analyze results of application processing 

```{r}
# Read in the Excel
application_results <- read_excel('application_results.xlsx')

head(application_results)
```
```{r}
approved_df = application_results[application_results$Approval_Status == "Approved",]
declined_df = application_results[application_results$Approval_Status == "Denied",]
```



```{r}
# ----------------  Results Analysis  -----------------

# Mean Line of Credit
mean_credit_line <- mean(application_results$Line_of_Credit, na.rm = T)
mean_credit_line
# Mean APR
mean_apr <- mean(application_results$APR, na.rm = T)
mean_apr
num_approved <- length(which(application_results$Approval_Status == "Approved"))
num_declined <- length(which(application_results$Approval_Status == "Denied"))

# How many applications were approved and percentage
pct_approved <- num_approved / (num_approved+num_declined)

# How many applications declined and percentage
pct_declined <- num_declined / (num_approved+num_declined)

```
* Analysis of our results show the following:
  + The average line of credit was $`r mean_credit_line`
  + The average APR was `r mean_apr`%
  + `r num_approved` (`r pct_approved`%) applicants were approved
  + `r num_declined` (`r pct_declined`%) applicants were declined

## Results Visualizations
```{r}

results_cloud_count <- count(approved_df, OCCUPATION_TYPE)

wordcloud2(data=results_cloud_count, size=1.6)

```

In the above word cloud, we see that the top applicants that were approved were Laborers, Sales Staff, and Core Staff. 

```{r}

approved_male <- length(which(approved_df$CODE_GENDER == "M"))
approved_female <- length(which(approved_df$CODE_GENDER == "F"))


approved_m_f_df <- data.frame(c("M","F"),c(approved_male,approved_female))
colnames(approved_m_f_df) <- c("Sex","Count")

# Male/Female Pie chart
ggplot(approved_m_f_df, aes(x="", y=Count, fill=Sex)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + ggtitle("Gender Approval Ratio")

approved_male_pct <- male/(approved_male+approved_female)
approved_female_pct <- female/(approved_male+approved_female)

```

```{r}
overall_approval_pct <- (approved_male + approved_female) / 438557
total_approvals <- approved_male + approved_female
```


The ratio of Male to Female, when observing our approvals, seems largly unchanged. 

```{r}
# Age box plot
ggplot(approved_df, aes(x=Age)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + ggtitle("Approval Age Distribution")

```

